---
title: "Creating and uploading Mapbox vector tilesets"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, warning = FALSE)
```

__mapboxapi__ includes a series of tools to help R users generate vector tiles from their data to be used in Mapbox Studio visualization projects.  The function `upload_tiles()` allows you to upload an sf object or a spatial file on disk to your Mapbox account using the Mapbox Uploads API. This will convert your spatial data into vector tiles that can be added to a style in Mapbox Studio.

Visualization of massive datasets, however, may require specialized tools.  Mapbox recommends the [tippecanoe](https://github.com/mapbox/tippecanoe) tile-generation utility for these tasks.  tippecanoe generates a scale-sensitive `.mbtiles` file that adds or removes complexity in a large dataset at different zoom levels.  

__mapboxapi__ allows R users to call tippecanoe from R with the `tippecanoe()` function.  This requires that tippecanoe is installed on your system; [see the tippecanoe documentation for installation instructions](https://github.com/mapbox/tippecanoe).  Once installed, tippecanoe can be used in large visualization workflows in concert with Mapbox Studio, such as the example below with Microsoft building footprints data for Texas.  

```{r, eval = FALSE}
library(mapboxapi)
library(mapdeck)
library(httr)

# Get the Microsoft buildings data for Texas and unzip
GET("https://usbuildingdata.blob.core.windows.net/usbuildings-v1-1/Texas.zip",
    write_disk("Texas.zip", overwrite = TRUE), progress())

unzip("Texas.zip")

# Use tippecanoe to make a dynamic .mbtiles file that visualizes large data appropriately
# at any zoom level.  sf objects can also be used as input!
# (requires installing tippecanoe on your machine separately first)
tippecanoe(input = "Texas.geojson",
           output = "Texas.mbtiles",
           layer_name = "texas_buildings")

# Upload the generated tileset to your Mapbox account (requires a Mapbox secret access token
# to be set as an environment variable)
upload_tiles(input = "Texas.mbtiles", username = "kwalkertcu",
             tileset_id = "TX_buildings",
             multipart = TRUE)


# Head over to Mapbox Studio when the upload is done (check the status with
# `check_upload_status()`) and add it to a style.  When you've styled it, bring it back
# into R with mapdeck by referencing the style ID:
mapdeck(token = Sys.getenv("MAPBOX_PUBLIC_TOKEN"),
        style = "mapbox://styles/kwalkertcu/ckaf9qxim1pyk1io7r2e8exj2/draft",
        zoom = 6,
        location = c(-98.7382803, 31.7678448))
```

<img src=img/tippecanoe-example.gif style="width: 800px">

These tools can also be used as part of larger research and visualization workflows, such as my [Mapping Immigrant America project](http://personal.tcu.edu/kylewalker/immigrant-america/#11/41.8643/-87.6307). 

<img src=img/immigrant-america.png style="width: 800px">


