---
title: "Mapping with mapboxapi"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, message = FALSE, warning = FALSE)
```

While __mapboxapi__ is not an interface to [Mapbox GL JS](https://docs.mapbox.com/mapbox-gl-js/api/), the Mapbox JavaScript library for web mapping, it does have functionality for map-making and visualization in combination with your spatial analysis workflows.  

## Interactive and static mapping with Mapbox styles

Let's say that you've designed a custom map style in Mapbox Studio that you'd like to incorporate into your R projects.  In this example, I'll use Mapbox's [Cartogram tool](https://apps.mapbox.com/cartogram/) to generate a custom map style from the R logo.  

<img src=img/cartogram.png style="width: 800px">

I can retrieve the style ID from my Mapbox account, or use the `list_styles()` function in __mapboxapi__ to get the ID.  With the style ID and my Mapbox username, the `addMapboxTiles()` function allows for direct incorporation of this style in an R Leaflet project.  

```{r}
library(leaflet)
library(mapboxapi)

leaflet() %>%
  addMapboxTiles(style_id = "ckb83x2ur02oa1hpiq9zdt1i2",
                 username = "kwalkertcu") %>%
  setView(lng = -74.0051,
          lat = 40.7251,
          zoom = 13)
```

In some cases, you may be more interested in a static map rather than an interactive one.  Static maps can be created with __mapboxapi__ in two ways.  Static maps can be generated directly in __mapboxapi__ with the `static_mapbox()` function.  This function accepts sf polygon/linestring (of limited size) or marker overlays.  

```{r}
points_of_interest <- tibble::tibble(
  longitude = c(-73.99405, -74.00616, -73.99577, -74.00761),
  latitude = c(40.72033, 40.72182, 40.71590, 40.71428)
  )

prepped_pois <- prep_overlay_markers(
  data = points_of_interest,
  marker_type = "pin-l",
  label = 1:4,
  color = "fff"
)

map <- static_mapbox(
  style_id = "ckb83x2ur02oa1hpiq9zdt1i2",
  username = "kwalkertcu",
  overlay_markers = prepped_pois, 
  width = 1200,
  height = 800
)

map
```

The output is an object from the [magick](https://github.com/ropensci/magick) R package that can be further customized with __magick__ functions.  

Alternatively, users might be interested in retrieving static tiles for use as a basemap in tmap or ggplot2 maps.  For an input sf object, the `get_static_tiles()` function will retrieve static tiles as a raster object around the input.  The example below fetches the standard Light Mapbox basemap and uses it as a basemap for demographic data obtained with tidycensus.  

```{r}
library(tidycensus)
library(tmap)

travis_income <- get_acs(
  geography = "tract",
  variables = "B19013_001",
  state = "TX",
  county = "Travis",
  geometry = TRUE,
  year = 2019
)

travis_tiles <- get_static_tiles(
  location = travis_income,
  zoom = 9,
  style_id = "light-v9",
  username = "mapbox"
)

tm_shape(travis_tiles) + 
  tm_rgb() + 
  tm_shape(travis_income) + 
  tm_polygons(col = "estimate",  
              alpha = 0.5, palette = "cividis", 
              title = "Median household income\n2015-2019 ACS") + 
  tm_layout(legend.outside = TRUE) + 
  tm_credits("Basemap © Mapbox, © OpenStreetMap", position = c("RIGHT", "BOTTOM"))
```


## Reading vector tile data into R

Mapbox styles on interactive maps are generally delivered as vector tiles, and can be visualized in this way in R with the [mapdeck R package](https://symbolixau.github.io/mapdeck/articles/mapdeck.html) by providing style information obtained from Mapbox Studio.  __mapboxapi__ also allows you to read data from vector tilesets directly into R as sf objects.  

For example, let's say I want to grab data from a vector tileset nearby to one of the points of interest I just mapped.  This can be accomplished with the `get_vector_tiles()` function and an appropriate tileset ID, which can be obtained from your Mapbox account.  The location to query can be a coordinate pair, an address, or an sf object. 

```{r}
vector_extract <- get_vector_tiles(
  tileset_id = "mapbox.mapbox-streets-v8",
  location = c(-73.99405, 40.72033),
  zoom = 15
)

names(vector_extract)
```
I get back a number of layers, including roads, labels, and buildings, from the vector tile at zoom 15 that intersects my input coordinates.  All of these elements are processed as sf objects that can be mapped with R tools:

```{r}
library(ggplot2)

ggplot(vector_extract$building$polygons) + 
  geom_sf() + 
  theme_void()
```

