---
title: "Using Mapbox navigation services with mapboxapi"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, warning = FALSE, message = FALSE)
options(tigris_use_cache = TRUE)
```

__mapboxapi__ supports the use of Mapbox's Directions, Isochrone, Matrix, and Optimization APIs, and are designed to be incorporated into R analysis workflows using sf, Shiny, and other packages. Navigation output should be visualized on a Mapbox map; see the vignette on mapping with __mapboxapi__ for how to accomplish this in R. 

## Routing with mapboxapi

The `mb_directions()` function computes a route between an origin and destination, or optionally along multiple points in an sf object.  Output options include the route or the route split by route legs as an sf linestring, or the full routing output as an R list for additional applications.  

```{r}
library(mapboxapi)
library(leaflet)

my_route <- mb_directions(
  origin = "10 Avenue de Wagram, 75008 Paris France",
  destination = "59 Rue de Tocqueville, 75017 Paris France",
  profile = "cycling",
  steps = TRUE,
  language = "fr"
)

leaflet(my_route) %>%
  addMapboxTiles(style_id = "light-v9",
                 username = "mapbox") %>%
  addPolylines()
```

Instructions are available in a variety of different languages:

```{r}
my_route$instruction
```
Optimized routes along multiple coordinates are available with the `mb_optimized_route()` function:  

```{r, fig.width = 8}
library(sf)

to_visit <- data.frame(
    X = c(-0.209307, -0.185875, -0.216877, -0.233511, -0.234541),
    Y = c(5.556019, 5.58031, 5.582528, 5.566771, 5.550209)
  ) %>%
  st_as_sf(coords = c("X", "Y"), crs = 4326)

optimized_route <- mb_optimized_route(to_visit,
                                      profile = "driving-traffic")

optimized_map <- static_mapbox(
  style_id = "light-v9",
  username = "mapbox",
  overlay_sf = optimized_route$route,
  overlay_style = list(
    stroke = "#144708"
  ),
  overlay_markers = prep_overlay_markers(
    data = optimized_route$waypoints,
    marker_type = "pin-l",
    label <- optimized_route$waypoints$waypoint_index,
    color = "144708"
  ),
  height = 800,
  width = 1200
)

optimized_map
```

## Isochrones and travel-time matrices

The Mapbox Isochrone API returns polygons representing _isochrones_, or areas reachable within a given travel time, around a given location.  __mapboxapi__ returns Mapbox isochrones as sf objects for visualization and analysis.  Up to 4 isochrone contours can be be requested for a given location with the `time` argument; input locations can be coordinate pairs, addresses, or sf objects.  

```{r}
library(mapdeck)

isochrones <- mb_isochrone("The Kremlin, Moscow Russia", 
                           time = c(4, 8, 12),
                           profile = "walking") 

mapdeck(style = mapdeck_style("light")) %>%
  add_polygon(data = isochrones, 
              fill_colour = "time",
              fill_opacity = 0.5,
              legend = TRUE) 
```

Travel-time matrices are available in the `mb_matrix()` function.  Requests are rate-limited internally to conform with Mapbox API limits; I've made internal modifications for larger-scale matrices but very large many-to-many matrices (more than 25 source coordinates and 25 destination coordinates) are not yet supported.  `mb_matrix()` returns an R matrix object where each cell represents the travel time between an origin and destination pair.  Coordinate pairs, addresses, and sf objects can be used as inputs.  If sf polygons are used as the input, `mb_matrix()` will calculate travel times based on the polygon centroids, which should be double-checked for possible routing errors.  

```{r}
library(tigris)
philly_tracts <- tracts("PA", "Philadelphia", cb = TRUE, class = "sf")
downtown_philly <- mb_geocode("Philadelphia City Hall, Philadelphia PA")

time_to_downtown <- mb_matrix(philly_tracts, downtown_philly)

philly_tracts$time <- time_to_downtown

mapdeck(style = mapdeck_style("light")) %>%
  add_polygon(data = philly_tracts, 
              fill_colour = "time",
              fill_opacity = 0.6,
              legend = TRUE)
```

